Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations

### 2022-11-01 paper/paper/SPIE/article.tex
[WIP]

With the increase and exapansion of digitial and electronic
Extracting useful features from complex data is a perennial challenge.

>>>>

\textbf{NOTE}: Forget academia for a moment, and write like its a blog. Then go back and add
the citations and ajust the tone.

Random Matrix Theory (RMT) has obvious potential. It has helped to understand the generalization
behaviour of neural networks [add citation], can be used to guide any principal components analysis
reduction [cite: Google search: "random matrix" "principal components"], has been used to help
understand phenomena as diverse as [include the usual citations to buses, traffic, physics, etc],
and can provide insights wherever there is a system of components interacting with some
independence [cite Cambridge RMT handbook, others].

At its most rudimentary, RMT simply describes the statistically-expected behaviour of the
eigenvalues of certain classes—or \textit{ensembles}—of random matrices under certain assumptions or
conditions. A random matrix is, generally, a matrix in which the entries can be considered to be
independently and identically distributed (iid), and where the iid distribution is a classic
mathemtical distribution. For example, one of the simplest ensembles [cite handbook] is the Gaussian
Unitary Ensemble (GUE), wherein [desc of ensemble properties] and which appears to be useful in
describing [real-world examples].

This means that if we examine a system and find that the distribution of eigenvalues of that system
resemble some RMT predicted spectrum, we potentially gain insight into the nature of that system.
If, in addition, we find a large deviance from the spectrum predicted by RMT, we find contrapositive
evidence that the system is \textit{not} like those described by RMT. Roughly, systems
well-described by RMT are highyl random and independent, whereas systems that are not so
well-described are either highly dependent, non-random, or both.

Despite the mathematical elegance and simplicity, RMT has yielded insights in variosu complex domains.
